SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-12 docs/questioning.tex
further fixes to Section VI

Other factors matter as well. Reporting absolute measurements gives
little information on the real overhead, unless separate measurements
are given for the system under study with and without the proposed
mechanism. Percentage measurements are normally better. Outliers may
be important. It is always necessary to describe the experimental
setup.

@@

\begin{figure*} [ht]
\begin{center}
\leavevmode
\includegraphics[scale=0.65]{defenses.pdf}
\end{center}
\caption{\label{fig:defenses}The basic categories of countermeasures
against web application attacks, based on source code
injectiion. For each approach we provide the references
that present corresponding mechanisms:
{\it Policy Enforcement}:~\cite{NSS06,JKK06a,KKVJ06,KJKV09,TNH07,RDWDE07,YCIS07,OWVS08,PSC09,ML10,DDHPJ10,PS11,VDDPJ11,BV08,LV09},
{\sc isr}:~\cite{BK04,JB07,GC09,APKLM10},
{\it Parse-Tree Validation}:~\cite{BWS05,SW06},
{\it Taint Tracking}:~\cite{HCF05,PB05,XBS06,NLC07,VFJKKV07,PMP11,SLMS14},
{\it Training}:~\cite{LLW02,HO05,HO06,HO05b,VMV05,JEP08,WPLKK09,MS09,MKS09,MKLS11},
{\it Hybrid}:~\cite{BV08,LV09,SMS13}.}
\end{figure*}

We categorize and describe the various mechanisms developed to detect
the attacks described in Section~\ref{sec:attacks} on the dimensions
we proposed apart from availability, which we treat separately in the
``Discussion" section (\ref{sec:discussion}).

Figure~\ref{fig:defenses} presents a taxonomy of the
defenses that protect web applications against attacks based on
code injection vulnerabilities.
We can identify three approaches:
{\it etiological}, {\it symptomatic} and {\it hybrid}.
The etiological category involves mechanisms designed to
stop attacks based on their causes and origins~\cite{JL75,L81}. 
The symptomatic category incorporates a variety of schemes that
inspect the behavior of applications and detect attacks based on
their undesirable symptoms~\cite{D76,A00}.
Hybrid mechanisms borrow characteristics from both
categories. Table~\ref{tab:comp} groups the various subcategories and
for every mechanism provides the following:

@@

Finally, as we mentioned in Subsection~\ref{sec:model},
Figure~\ref{fig:attacks} illustrates for each mechanism (found in the
figure's caption), the point where it detects attacks.

@@

The key idea behind this approach is to compare
the tree representation of the abstract syntactic
structure of the source code that is about to be
executed with the one that was originally intended.
If the trees diverge, the application is probably
under attack.

In the case of {\sc dsl} injection attacks, mechanisms check 
the query before the inclusion of user input with the one
resulting after the inclusion of user input.
Both mechanisms that implement this approach,
{\sc sqlg}uard~\cite{BWS05} and
{\sc sql}check~\cite{SW06} are pretty similar
and detect the attack before a query reaches the
database (Point {\sc e}i{\sc q}).
Contrary to {\sc sqlg}uard (and other
mechanisms as we will observe), {\sc sql}check has been
extensively tested in terms of accuracy
as it can be seen in Table~\ref{tab:comp}.
A disadvantage of those mechanisms is that the
application must be modified in every code fragment
that sends an {\sc sql} query to the database for execution.

Parse-tree validation seems to be an efficient approach
to detect {\sc dsl} code injection attack,
especially in the case of the {\sc sql}check
implementation. This however does not apply in
the case of the mechanisms that borrow elements from
this approach and examine the syntax trees
of scripts to detect JavaScript-driven {\sc xss} attacks
as we will see in Subsection~\ref{sec:hybrid}.

@@

This approach is used to detect {\sc xss} and {\sc csrf} attacks.
When using a framework that implements
this approach, developers must define
specific security policies on the server side.
Policies can be expressed through JavaScript extensions,
pattern matching or syntax specific settings.
Then the policies are enforced either in the user's
browser at runtime or in a server-side proxy that intercepts
server responses.

{\it Noxes}~\cite{KKVJ06,KJKV09} is the only
framework that partially allows the web
user to specify policies to prevent {\sc xss} attacks.
The key idea behind Noxes is to parse
the {\sc html} that reaches the browser to
find static {\sc url} references. Then, based on
a set of policies Noxes allows or blocks
the request (Point {\sc ub}). Such policies
can be also provided by the server (i.e. ``never
follow a link that leads to the malicious.com
web site"). The main issue with Noxes is that
it does not deal with JavaScript-driven {\sc xss}
attacks which is currently the most popular
{\sc xss} attack vector.

A number of the frameworks define policies
based on information and features provided by
the Document Object Model ({\sc dom}) of a web page.
Specifically, developers must place all legitimate
scripts inside {\sc html} elements like {\tt div}.
The web browser (Point {\sc ub}), parses the
{\sc dom} tree and executes scripts only when they
are contained in such elements. All other scripts
are treated according to the policies defined
on the server. Frameworks that support this
functionality are {\sc beep}~\cite{TNH07}
and {\sc dsi}~\cite{NSS06}. The main problem with
these mechanisms is that they do
not examine the script's location inside the web
document. Based on this fact, attackers
can perform mimicry attacks~\cite{WS02}.
Specifically, they can execute legitimate scripts,
but not as intended by the original design of
the developers. This is extensively described
in reference~\cite{APKLM10}. In addition,
the authors of~\cite{APKLM10} describe
another recent variation of JavaScript
injection attacks that can be used to bypass the
above mechanisms, known as return-to-JavaScript
attacks.
% Such attacks share
% many similarities with the {\it return-to-lib}
% attack~\cite{SPPGMMB04}. 
%% PL Explain what it does, I guess it's what you go on to say
%% directly below?
% When performing an attack like this the malicious user transfers the
% execution to points that contain script code which in turn is used to
% perform the attack.

Another policy enforcement approach introduces
policies directly either in {\sc html} or JavaScript code
to confine their behavior. {\it BrowserShield}~\cite{RDWDE07}
acts as a proxy on the server side (Point {\sc s}t{\sc b}) to
parse the {\sc html} of server responses and identifies
scripts. Then, it rewrites them into safe equivalents
and protects the web user from exploits
that are based on reported browser vulnerabilities.
{\it ConScript}~\cite{ML10}, {\it CoreScript}~\cite{YCIS07}
and the framework by {\it Phung et al.}~\cite{PSC09}
extend JavaScript with new primitive functions that
provide safe methods to protect potentially vulnerable
JavaScript functions. In both cases, policy enforcement takes
place on the JavaScript engine of the browser (Point {\sc ub}).
In this way, {\sc xxs} attacks that take advantage
functions like {\tt write} and {\tt eval}, which are
used to assemble innocuous-looking parts into harmful
strings,\footnote{The infamous Sammy worm that
infected MySpace in 2005~\cite{SP07,ELX07}
utilized the {\tt eval} function to assemble a
script with malicious behavior.} would fail.

@@

\noindent
{\it WebJail}~\cite{VDDPJ11}, and
{\sc soma}~\cite{OWVS08} are two frameworks that
can actually detect such an attack (Point {\sc ub}).
To achieve this, {\sc soma} requires site administrators
to specify legitimate, external domains for sending
or receiving information in order to approve interactions
between them and the protected web site. As a result,
{\sc soma} can also detect {\sc csrf} attacks.
WebJail, contains the functionality of
third-party scripts by introducing a web component
integrator that restricts the access that these scripts
may have to either the data or the functionality of other components

Policy enforcement mechanisms that detect {\sc csrf}
attacks are usually implemented in the form of a
server-side proxy (Point {\sc s}t{\sc b})
interposing on the client-server communication, and modifying
it. {\it NoForge}~\cite{JKK06a}
parses the {\sc html} server responses
and adds a token to every {\sc url} referring to this
server. Then it associates the token with the cookie
representing the session {\sc id} for the application.
When a request is received, the mechanism checks
if the request contains the token related
to the session {\sc id}. A disadvantage of NoForge
is that {\sc html} that is dynamically created within
the browser will not include the token.
Thus, sites that create their {\sc html} on the client
will remain vulnerable. In addition, it does not
support cross-origin requests.
{\it j{\sc csrf}}~\cite{PS11} has a similar
functionality and meanwhile addresses the above problems.
Finally, {\it CsFire}~\cite{DDHPJ10}
examines cross-domain interactions, to design a
cross-domain policy on the client's side (Point {\sc ub}).
The policy is based on the concept of a relaxed
same-origin policy that allows communication between
sub-domains of the same registered domain.

@@

{\sc isr} is a method that has been applied to counter different kinds
of application attacks~\cite{K09b,KKP03}. The main idea behind it is
to separate code from data by randomizing the legitimate code's
execution environment. In this way, the malicious code injected by the
attackers who do not know the randomization algorithm, will not get
executed.

{\it {\sc sql}rand}~\cite{BK04} implements {\sc isr} to detect {\sc
  sql} injection attacks. It allows programmers to create {\sc sql}
statements using randomized instructions instead of standard keywords.
The modified queries are reconstructed at runtime using the same key
that is inaccessible to the malicious user. {\sc sql}rand is one of
the few mechanisms that detect {\sc sql} injection attacks on the
database level (Point i{\sc db}).

In the case of {\sc xss} attacks that are based on JavaScript or {\sc
  html}, consider a {\sc xor} function that encodes all source of a
web page on the server-side and then, on the client-side, the web
browser decodes the source by applying the same function again (Point
{\sc ub}). Variations of this approach include {\it
  Noncespaces}~\cite{GC09} and {\it x{\sc js}}~\cite{APKLM10}, which
randomize the instruction set of {\sc html} and JavaScript
respectively. Contrary to x{\sc js}, in Noncespaces administrators
must set specific policies in a manner similar to a firewall
configuration language. {\it {\sc sm}ask}~\cite{JB07} is another
framework that was inspired by {\sc isr}. To detect {\sc xss}
attacks it searches for {\sc html} and
JavaScript keywords within the application's legitimate code. This is
done before the processing of an {\sc http} request. When it finds one
it adds a token to it. This results to a ``code mask". Then, before
sending the resulting {\sc html} data to the web user, the framework
searches the data for illegal code by using the same keywords (Point
{\sc s}t{\sc b}). Since all legitimate code has been ``masked", the injected
code can be identified. The pre- and post-processing of the code
though, may add a significant overhead to the application.
Unfortunately, the authors of {\sc sm}ask did not provide measurements
regarding the computational performance of the tool (see
Table~\ref{tab:comp}).

{\sc isr} is a deterministic approach that can be applied to detect
different attacks in an efficient manner. However, Sovarel et
al.~\cite{SEP05} have investigated thoroughly the effectiveness of
{\sc isr} and showed that a malicious user may be able to circumvent
it by determining the randomization key and their results indicate
that doing {\sc isr} in a way that provides a certain degree of
security against a motivated attacker is more difficult than
previously thought. Furthermore, developers who wish to use such
mechanisms must follow good coding practices and never show a
randomized code statement in any case (i.e. an exception error)
because in this way they will reveal the encoding key.

Even though the above implementations impose a low computational
overhead, they impose an infrastructure overhead. In particular, in
{\sc sql}rand requires the integration of a proxy within the database
server, while Noncespaces and x{\sc js} require modifications on both
the server and the client.

@@

In this section we examine the two main symptomatic
approaches.

@@

A taint tracking scheme marks untrusted
(``tainted") data (for instance, a variable set
by a field in a web form) and traces its flow through
the program. If the variable is used in an expression
that sets another variable, that variable is also
marked as untrusted and so on. If any of these variables
is used to execute a potentially risky command (``sink") the
scheme may act accordingly.

@@

There are different implementations of this approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (Point {\sc e}i{\sc q}).

It is possible to apply further checks when it is established that
tainted data have reached a sink (Point {\sc e}i{\sc q}). {\it Xu et
  al.}~\cite{XBS06} track taint information at the level of bytes in
memory. To distinguish between legitimate and malicious uses of
untrusted data that reach a sink, they search the data for suspicious
symbols by using regular expressions. {\sc csse}~\cite{PB05}
associates tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application,
and others. When tainted data reaches a sink, {\sc csse} performs
syntactic checks based on the aforementioned metadata (Point {\sc e}i{\sc q}).
{\it {\sc php} Aspis}~\cite{PMP11} works in a similar way. To obtain
metadata, it takes advantage of {\sc php} arrays data structure.
Finally, {\sc wasc}~\cite{NLC07} analyzes {\sc html} responses to check
if there are tainted data that contain scripts (Point {\sc s}t{\sc b}).

A recent study~\cite{NBR14} showed that there are ways to circumvent
the majority of the above schemes. Furthermore, most of them are not
easy to deploy since the majority of input vectors, string operations
and output vectors of the application must be instrumented.

@@

Training techniques are based on the ideas of Denning's original
intrusion detection framework~\cite{Den87}. In particular, a training
mechanism registers all valid legitimate code statements during a
training phase (mostly in the form of signatures). This can
be done in various ways according to the implementation. Then, only
those will be recognized and approved for execution
during production.

Training methods that detect {\sc dsl}-driven injection attacks
generate and store valid code statements (i.e. {\sc sql} or {\sc
  xp}ath queries) in various forms, and detect attacks as outliers
from the set of valid code statements. An early approach, {\it {\sc
    didafit}}~\cite{LLW02} detects {\sc sql} injection (Point
sdb) attacks by recording all database transactions
(stripped from user input). Subsequent
refinements by Valeur et al.~\cite{VMV05} tagged each transaction with
the corresponding application as an extension of their anomaly
detection framework called {\it libAnomaly}. {\it {\sc
    sd}river}~\cite{MS09,MKS09,MKLS11} is a signature-based mechanism
that prevents {\sc sql} and {\sc xp}ath injection attacks. The
signatures generated during a training phase are based on features
that can depend either on the code statement or on its execution
environment (i.e. the stack trace). Then, at runtime, the mechanism
checks all statements for compliance and can block code statements
containing injected elements (Point {\sc e}i{\sc q}). {\sc
  amnesia}~\cite{HO05,HO06,HO05b} is a tool that also detects {\sc
  sql} injection attacks (Point {\sc e}i{\sc q}) by associating a query model
with the location of every {\sc sql} statement within the application.
Then, at runtime, it monitors the application's execution to detect
when {\sc sql} statements diverge from the expected model.

Various countermeasures against {\sc xss} attacks follow a similar
pattern. {\it {\sc swap}}~\cite{WPLKK09} creates a unique identifier
({\it script {\sc id}}) for every legitimate script on the server.
Then, a JavaScript detection component placed in a web proxy (Point
{\sc s}t{\sc b}) searches for injected scripts with no corresponding {\sc id}
in the server's responses. If no injected scripts are found, the proxy
forwards the request to the client. This mechanism is relatively
inflexible since it does not support dynamic scripts. In addition it
imposes a significant overhead (see Table~\ref{tab:comp}). The authors
of {\it {\sc xssds}}~\cite{JEP08} have implemented a similar mechanism
that also supports dynamic and external scripts. Specifically, during
training phase they build a list of all benign scripts. In the case of
external scripts they keep a white list of all the valid domain names
that contain scripts used by the application.

Defenses based on training include some mechanisms that can be easily
circumvented. For example, {\sc didafit} and {\it libAnomaly}
do not tag transactions with their corresponding call-sites.
In this case a false negative result
is prominent. For instance, consider an application that will send the
password for a forgetful user via email by executing the following
query:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
In this case {\sc didafit} and {\it libAnomaly}, could be easily escaped
since a malicious user, could obtain the administrator's password
by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent

However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
Furthermore, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect mimicry attacks 
where legitimate scripts can be executed by
attackers, but not in the way intended by
the the developers~\cite{APKLM10}.

In general, the detection accuracy of training approaches is
heavily based on the coverage that is achieved during the
training phase. If the coverage is insufficient the
production of false alarms is very likely.
In addition, when a code statement is altered, a new
training phase is necessary.

Most of the mechanisms above are relatively easy to deploy.
{\sc dsl} injection attack countermeasures
can be retrofitted to a system typically by changing
some configuration files (i.e. {\sc sd}river). This does not apply
to {\sc amnesia} though, since significant source code
modifications are required for every query that exists
in the application. Finally, {\sc swap} and {\sc xssds}
are implemented within a proxy on the server-side.

@@

This category includes mechanisms that borrow
characteristics from more than one of the above categories.
Two of them focus on the detection of {\sc xss}
attacks and one of them focuses on {\sc dsl} code
injection attacks.

>>>>

What kind of numbers is reported matters as well. Reporting absolute
measurements gives little information on the real overhead, unless
separate measurements are given for the system under study with and
without the proposed mechanism. Percentage measurements are normally
better.

@@

\begin{figure*} [ht]
\begin{center}
\leavevmode
\includegraphics[scale=0.65]{defenses.pdf}
\end{center}
\caption{\label{fig:defenses}The basic categories of countermeasures
against web application attacks, based on source code
injection. For each approach we provide the references
that present corresponding mechanisms:
{\it Policy Enforcement}:~\cite{NSS06,JKK06a,KKVJ06,KJKV09,TNH07,RDWDE07,YCIS07,OWVS08,PSC09,ML10,DDHPJ10,PS11,VDDPJ11,BV08,LV09},
{\sc isr}:~\cite{BK04,JB07,GC09,APKLM10},
{\it Parse-Tree Validation}:~\cite{BWS05,SW06},
{\it Taint Tracking}:~\cite{HCF05,PB05,XBS06,NLC07,VFJKKV07,PMP11,SLMS14},
{\it Training}:~\cite{LLW02,HO05,HO06,HO05b,VMV05,JEP08,WPLKK09,MS09,MKS09,MKLS11},
{\it Hybrid}:~\cite{BV08,LV09,SMS13}.}
\end{figure*}

We categorize and describe the various mechanisms developed to detect
the attacks described in Section~\ref{sec:attacks} on the dimensions
we proposed apart from availability, which we treat separately in the
discussion in Section (\ref{sec:discussion}).

Figure~\ref{fig:defenses} presents a taxonomy of the
defenses that protect web applications against attacks based on
code injection vulnerabilities.
We can identify three approaches:
{\it etiological}, {\it symptomatic}, and {\it hybrid}.
The etiological category involves mechanisms designed to
stop attacks based on their causes and origins~\cite{JL75,L81}. 
The symptomatic category incorporates a variety of schemes that
inspect the behavior of applications and detect attacks based on
their undesirable symptoms~\cite{D76,A00}.
Hybrid mechanisms borrow characteristics from both
categories. Table~\ref{tab:comp} groups the various subcategories and
for every mechanism provides the following:

@@

Recall that the point where attacks are detected is given in the
caption of Figure~\ref{fig:attacks}.

@@

The key idea behind parse-tree validation is to compare the tree
representation of the abstract syntactic structure of the source code
that is about to be executed with the one that was originally
intended. If the trees diverge, the application is probably under
attack.

In the case of {\sc dsl} injection attacks, mechanisms check 
the query before the inclusion of user input with the one
resulting after the inclusion of user input.
Both mechanisms that implement this approach,
{\sc sqlg}uard~\cite{BWS05} and
{\sc sql}check~\cite{SW06} are pretty similar
and detect the attack before a query reaches the
database (Point {\sc e}i{\sc q}).
Contrary to {\sc sqlg}uard and other
mechanisms, {\sc sql}check has been
extensively tested in terms of accuracy
as can be seen in Table~\ref{tab:comp}.
A disadvantage of those mechanisms is that the
application must be modified in every code fragment
that sends an {\sc sql} query to the database for execution.

Parse-tree validation seems to be an efficient approach
to detect {\sc dsl} code injection attack,
especially in the case of the {\sc sql}check
implementation. This however does not apply in
the case of the mechanisms that borrow elements from
this approach and examine the syntax trees
of scripts to detect JavaScript-driven {\sc xss} attacks,
as we will see in Subsection~\ref{sec:hybrid}.

@@

This approach is used to detect {\sc xss} and {\sc csrf} attacks. When
using a framework that implements policy enforcement, developers must
define specific security policies on the server side. Policies can be
expressed through JavaScript extensions, pattern matching, or syntax
specific settings. Then the policies are enforced either in the user's
browser at runtime or in a server-side proxy that intercepts server
responses.

{\it Noxes}~\cite{KKVJ06,KJKV09} is the only
framework that partially allows the web
user to specify policies to prevent {\sc xss} attacks.
The key idea behind Noxes is to parse
the {\sc html} that reaches the browser to
find static {\sc url} references. Then, based on
a set of policies Noxes allows or blocks
the request (Point {\sc ub}). Such policies
can be also provided by the server (i.e. ``never
follow a link that leads to the malicious.com
web site"). The main issue with Noxes is that
it does not deal with JavaScript-driven {\sc xss}
attacks, which currently make for the most popular
{\sc xss} attack vector.

A number of the frameworks define policies based on information and
features provided by the Document Object Model ({\sc dom}) of a web
page. Specifically, developers must place all legitimate scripts
inside {\sc html} elements like {\tt div}. The web browser (Point {\sc
  ub}), parses the {\sc dom} tree and executes scripts only when they
are contained in such elements. All other scripts are treated
according to the policies defined on the server. Frameworks that
support this functionality are {\sc beep}~\cite{TNH07} and {\sc
  dsi}~\cite{NSS06}. The main problem with these mechanisms is that
they do not examine the script's location inside the web document.
Based on this fact, attackers can perform mimicry attacks~\cite{WS02}.
Specifically, they can execute legitimate scripts, but not as intended
by the original design of the developers. This is extensively
described by {\it Athanasopoulos et al.}~\cite{APKLM10}, who also describe
another recent variation of JavaScript injection attacks, known as
return-to-JavaScript attacks that can be used to bypass the above
mechanisms.
% Such attacks share
% many similarities with the {\it return-to-lib}
% attack~\cite{SPPGMMB04}. 
%% PL Explain what it does, I guess it's what you go on to say
%% directly below?
% When performing an attack like this the malicious user transfers the
% execution to points that contain script code which in turn is used to
% perform the attack.

Another policy enforcement approach introduces
policies directly either in {\sc html} or JavaScript code
to confine their behavior. {\it BrowserShield}~\cite{RDWDE07}
acts as a proxy on the server side (Point {\sc s}t{\sc b}) to
parse the {\sc html} of server responses and identify
scripts. Then, it rewrites them into safe equivalents
and protects the web user from exploits
that are based on reported browser vulnerabilities.
{\it ConScript}~\cite{ML10}, {\it CoreScript}~\cite{YCIS07}
and the framework by {\it Phung et al.}~\cite{PSC09}
extend JavaScript with new primitive functions that
provide safe methods to protect potentially vulnerable
JavaScript functions. In both cases, policy enforcement takes
place on the JavaScript engine of the browser (Point {\sc ub}).
In this way, {\sc xxs} attacks that take advantage
functions like {\tt write} and {\tt eval}, which are
used to assemble innocuous-looking parts into harmful
strings,\footnote{The infamous Sammy worm that
infected MySpace in 2005~\cite{SP07,ELX07}
utilized the {\tt eval} function to assemble a
script with malicious behavior.} would fail.

@@

\noindent {\it WebJail}~\cite{VDDPJ11}, and {\sc soma}~\cite{OWVS08}
are two frameworks that can actually detect such an attack (Point {\sc
  ub}). To achieve this, {\sc soma} requires site administrators to
specify legitimate, external domains for sending or receiving
information in order to approve interactions between them and the
protected web site. As a result, {\sc soma} can also detect {\sc csrf}
attacks. WebJail contains the functionality of third-party scripts by
introducing a web component integrator that restricts the access that
these scripts may have to either the data or the functionality of
other components.

Policy enforcement mechanisms that detect {\sc csrf}
attacks are usually implemented in the form of a
server-side proxy (Point {\sc s}t{\sc b})
interposed on the client-server communication and modifying
it. {\it NoForge}~\cite{JKK06a}
parses the {\sc html} server responses
and adds a token to every {\sc url} referring to this
server. Then it associates the token with the cookie
representing the session {\sc id} for the application.
When a request is received, the mechanism checks
if the request contains the token related
to the session {\sc id}. A disadvantage of NoForge
is that {\sc html} that is dynamically created within
the browser will not include the token.
Thus, sites that create their {\sc html} on the client
will remain vulnerable. In addition, it does not
support cross-origin requests.
{\it j{\sc csrf}}~\cite{PS11} has a similar
functionality and meanwhile addresses the above problems.
Finally, {\it CsFire}~\cite{DDHPJ10}
examines cross-domain interactions, to design a
cross-domain policy on the client's side (Point {\sc ub}).
The policy is based on the concept of a relaxed
same-origin policy that allows communication between
sub-domains of the same registered domain.

@@

{\sc isr} is a method that has been applied to counter different kinds
of application attacks~\cite{K09b,KKP03}. The main idea behind it is
to separate code from data by randomizing the legitimate code's
execution environment. In this way, the malicious code injected by the
attackers who do not know the randomization algorithm will not get
executed.

{\it {\sc sql}rand}~\cite{BK04} implements {\sc isr} to detect {\sc
  sql} injection attacks. It allows programmers to create {\sc sql}
statements using randomized instructions instead of standard keywords.
The modified queries are reconstructed at runtime using the same key
used for randomization, which is inaccessible to the malicious user.
{\sc sql}rand is one of the few mechanisms that detect {\sc sql}
injection attacks on the database level (Point i{\sc db}).

In the case of {\sc xss} attacks that are based on JavaScript or {\sc
  html}, consider a {\sc xor} function that encodes all source of a
web page on the server-side and then, on the client-side, the web
browser decodes the source by applying the same function again (Point
{\sc ub}). Variations of this approach include {\it
  Noncespaces}~\cite{GC09} and {\it x{\sc js}}~\cite{APKLM10}, which
randomize the instruction set of {\sc html} and JavaScript
respectively. Contrary to x{\sc js}, in Noncespaces administrators
must set specific policies in a manner similar to a firewall
configuration language. {\it {\sc sm}ask}~\cite{JB07} is another
framework that was inspired by {\sc isr}. To detect {\sc xss}
attacks it searches for {\sc html} and
JavaScript keywords within the application's legitimate code. This is
done before the processing of an {\sc http} request. When it finds one
it adds a token to it. This results to a ``code mask". Then, before
sending the resulting {\sc html} data to the web user, the framework
searches the data for illegal code by using the same keywords (Point
{\sc s}t{\sc b}). Since all legitimate code has been ``masked", the injected
code can be identified. The pre- and post-processing of the code,
though, may add a significant overhead to the application.
Unfortunately, the authors of {\sc sm}ask did not provide measurements
regarding the computational performance of the tool (see
Table~\ref{tab:comp}).

{\sc isr} is a deterministic approach that can be applied to detect
different attacks in an efficient manner. However, {\it Sovarel et
al.}~\cite{SEP05} have investigated thoroughly the effectiveness of
{\sc isr} and showed that a malicious user may be able to circumvent
it by determining the randomization key. Their results indicate
that doing {\sc isr} in a way that provides a certain degree of
security against a motivated attacker is more difficult than
previously thought. Furthermore, developers who wish to use such
mechanisms must follow good coding practices and never show a
randomized code statement in any case (i.e., an exception error)
because in this way they will reveal the encoding key.

Even though the above implementations impose a low computational
overhead, they impose a deployment infrastructure overhead. In particular, in
{\sc sql}rand requires the integration of a proxy within the database
server, while Noncespaces and x{\sc js} require modifications on both
the server and the client.

@@

Symptomatic approaches adopt one of two tacks. They either track
untrusted input and ban certain operations on it, or they first learn what
code to trust and then approve for execution code that they recognize
as safe.

@@

A taint tracking scheme marks untrusted (``tainted") data (for
instance, a variable set by a field in a web form) and traces its flow
through the program. If the variable is used in an expression that
sets another variable, that variable is also marked as untrusted and
so on. If any of these variables is used in a potentially risky
command (for example, send the data to a vulnerable ``sink'', such as
a database, a file, the network) the scheme may act accordingly.

@@

There are different implementations of this approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
an exception is raised (Point {\sc e}i{\sc q}).

It is possible to apply further checks when it is established that
tainted data have reached a sink (Point {\sc e}i{\sc q}). {\it Xu et
  al.}~\cite{XBS06} track taint information at the level of bytes in
memory. To distinguish between legitimate and malicious uses of
untrusted data that reach a sink, they search the data for suspicious
symbols by using regular expressions. {\sc csse}~\cite{PB05}
associates tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application,
and others. When tainted data reaches a sink, {\sc csse} performs
syntactic checks based on the aforementioned metadata (Point {\sc e}i{\sc q}).
{\it {\sc php} Aspis}~\cite{PMP11} works in a similar way. To obtain
metadata, it takes advantage of the  {\sc php} array data structure.
Finally, {\sc wasc}~\cite{NLC07} analyzes {\sc html} responses to check
if there are tainted data that contain scripts (Point {\sc s}t{\sc b}).

A recent study~\cite{NBR14} showed that there are ways to circumvent
the majority of the above schemes. Furthermore, most of them are not
easy to deploy since the majority of input vectors, string operations,
and output vectors of the application must be instrumented.

@@

Training techniques are based on the ideas of Denning's original
intrusion detection framework~\cite{Den87}. In particular, a training
mechanism registers all valid legitimate code statements during a
training phase (mostly in the form of signatures). This can be done in
various ways according to the implementation. Then, only those
statements will be recognized and approved for execution during
production.

Training methods that detect {\sc dsl}-driven injection attacks
generate and store valid code statements (i.e., {\sc sql} or {\sc
  xp}ath queries) in various forms, and detect attacks as outliers
from the set of valid code statements. An early approach, {\it {\sc
    didafit}}~\cite{LLW02} detects {\sc sql} injection (Point
{\sc sdb}) attacks by recording all database transactions
stripped from user input. Subsequent
refinements by Valeur et al.~\cite{VMV05} tagged each transaction with
the corresponding application as an extension of their anomaly
detection framework called {\it libAnomaly}. {\it {\sc
    sd}river}~\cite{MS09,MKS09,MKLS11} is a signature-based mechanism
that prevents {\sc sql} and {\sc xp}ath injection attacks. The
signatures generated during a training phase are based on features
that can depend either on the code statement or on its execution
environment (i.e., the stack trace). Then, at runtime, the mechanism
checks all statements for compliance and can block code statements
containing injected elements (Point {\sc e}i{\sc q}). {\sc
  amnesia}~\cite{HO05,HO06,HO05b} is a tool that also detects {\sc
  sql} injection attacks (Point {\sc e}i{\sc q}) by associating a query model
with the location of every {\sc sql} statement within the application.
Then, at runtime, it monitors the application's execution to detect
when {\sc sql} statements diverge from the expected model.

Various countermeasures against {\sc xss} attacks follow a similar
pattern. {\it {\sc swap}}~\cite{WPLKK09} creates a unique identifier
({\it script {\sc id}}) for every legitimate script on the server.
Then, a JavaScript detection component placed in a web proxy (Point
{\sc s}t{\sc b}) searches for injected scripts with no corresponding {\sc id}
in the server's responses. If no injected scripts are found, the proxy
forwards the request to the client. This mechanism is relatively
inflexible since it does not support dynamic scripts. In addition it
imposes a significant overhead (see Table~\ref{tab:comp}). The authors
of {\it {\sc xssds}}~\cite{JEP08} have implemented a similar mechanism
that also supports dynamic and external scripts. Specifically, during the
training phase they build a list of all benign scripts. In the case of
external scripts they keep a white list of all the valid domain names
that contain scripts used by the application.

Defenses based on training include some mechanisms that can be easily
circumvented. For example, {\sc didafit} and {\it libAnomaly} do not
tag transactions with their corresponding call sites. This can lead
easily to false negatives. For instance, consider an application that
will send the password for a forgetful user via email by executing the
following query: \bgroup \lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
In this case {\sc didafit} and {\it libAnomaly}, could be easily bypassed
since a malicious user could obtain the administrator's password
by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent
The infected query matches the signature of the second one above and
is therefore accepted. The problem lies with the call location of the
query, not the query alone.

Mechanisms that create signatures that involve elements not only
associated with the code statements (i.e., {\sc amnesia} and {\sc
  sd}river) could detect such attacks. Specifically, {\sc sd}river
associates a complete stack trace with the root of an {\sc sql}
statement, thus it can correlate queries with their call sites and
detect attacks like the aforementioned one. Furthermore, training
tools that detect {\sc xss} attacks based on JavaScript injection
would fail to detect mimicry attacks where legitimate scripts can be
executed by attackers, but not in the way intended by the the
developers~\cite{APKLM10}.

In general, the detection accuracy of training approaches is heavily
based on the coverage that is achieved during the training phase. If
the coverage is insufficient false alarms are very likely. In
addition, when a code statement is altered, a new training phase is
necessary.

Most of the training approaches are relatively easy to deploy.
{\sc dsl} injection attack countermeasures
can be retrofitted to a system typically by changing
some configuration files (i.e. {\sc sd}river). This does not apply
to {\sc amnesia} though, since significant source code
modifications are required for every query that exists
in the application. Finally, {\sc swap} and {\sc xssds}
are implemented within a proxy on the server-side.

@@

This category includes mechanisms that borrow
characteristics from both of the above categories.
Two of them focus on the detection of {\sc xss}
attacks and one of them focuses on {\sc dsl} code
injection attacks.
